---
title: "Probability Analysis"
author: "Abhay Paidipalli"
date: "6/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 9, fig.height = 5)

library(tidyverse)
library(ggthemes)
library(plotly)
library(matrixStats)
library(S4Vectors)
library(memoise)

```

## Weight Example

```{r, include=TRUE, echo=FALSE}
# contains all of our functions in order to run a simulation for the weight example

# gets sample from a uniformly distributed population
uniform_sample <- function(numPeople, lowerBound, upperBound) {
  result <- sample(lowerBound:upperBound, numPeople, replace=T)
}

# gets sample from a normally distributed population
normal_sample <- function(numPeople, mean, std) {
  result <- rnorm(numPeople, mean, std)
}

# gets sample from a geometrically distributed population
geometric_sample <- function(numPeople, prob) {
  result <- rgeom(numPeople, prob)
}

# creates resolution based on parameters
create_res <- function(numRes, lowerBound, upperBound) {
  extra_res <- numRes - 2
  if (extra_res > 0) {
    result <- c(lowerBound, upperBound)
    gap <- (upperBound - lowerBound) / (extra_res + 1)
    tempNum <- lowerBound
    for (i in 1:extra_res) {
      tempNum <- tempNum + gap
      signif(tempNum, digits=2)
      result <- append(result, tempNum)
    }
  } else {
    result <- c(lowerBound, upperBound)
  }
  return(result)
}

# applies the resolution to a sample
apply_res <- function(nums, res) {
  result <- c()
  for (num in nums) {
    shortest_distance <- max(nums) - min(nums)
    closest_res <- 0
    for (r in res) {
      if (abs(num - r) < shortest_distance) {
        shortest_distance <- abs(num - r)
        closest_res <- r
      }
    }
    result <- append(result, closest_res)
  }
  return(result)
}

# will do one comparison 
conduct_trial <- function(numGroups, numPeoplePerGroup, lowerBound, upperBound, numRes, sampleType) {
  avg <- c()
  for (i in 1:numGroups) {
    if (sampleType == "normal") {
      # using the rule that 3 standard deviations away from the mean equals 99.7% of the area
      sample <- normal_sample(numPeoplePerGroup, mean(c(lowerBound, upperBound)), (upperBound - lowerBound) / 3)
    } else if (sampleType == "uniform") {
      sample <- uniform_sample(numPeoplePerGroup, lowerBound, upperBound)
    } else if (sampleType == "geometric") {
      # using the rule that mean of geometric distribution is reciprocal of probability (will estimate this around a quarter of the way from lower to upper bounds)
      sample <- geometric_sample(numPeoplePerGroup, 1 / (mean(c(lowerBound, mean(c(lowerBound, upperBound))))))
    }
    resolution <- create_res(numRes, lowerBound, upperBound)
    sample_with_resolution <- apply_res(sample, resolution)
    avg <- append(avg, mean(sample_with_resolution))
  }
  return(min(avg) == max(avg))
}

# will do the entire simulation for a certain number of trials
conduct_sim <- function(numTrials, numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, numRes, sampleType) {
  result <- c()
  for (i in 1:numTrials) {
    result <- append(result, conduct_trial(numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, numRes, sampleType))
  }
  return(result)
}

# calculates ultimate probability that resulted for conditions of the simulation
calc_prob <- function(bools) {
  tru_bool_count <- 0
  fal_bool_count <- 0
  for (bool in bools) {
    if (bool == TRUE) {
      tru_bool_count <- tru_bool_count + 1
    } else {
      fal_bool_count <- fal_bool_count + 1
    }
  }
  return(tru_bool_count/length(bools))
}

# makes tibble using results of the prob function (resolution vs. probability)
create_prob_dist <- function(numTrials, numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, lowerRes, upperRes, concentration, sampleType) {
  prob_data <- tibble(numRes = numeric(), prob = numeric())
  numRes <- c()
  prob <- c()
  for (i in lowerRes:upperRes) {
    for (x in 1:concentration) {
      numRes <- append(numRes, i)
      prob <- append(prob, calc_prob(conduct_sim(numTrials, numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, i, sampleType)))
    }
  }
  prob_data %>% 
    add_row(numRes, prob)
}

# creates the graph: resolution vs. probability
create_prob_graph <- function(numTrials, numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, lowerRes, upperRes, concentration, sampleType) {
  data <- create_prob_dist(numTrials, numGroupsPerTrial, numPeoplePerGroup, lowerBound, upperBound, lowerRes, upperRes, concentration, sampleType) %>% 
    ggplot(mapping = aes(numRes, prob)) + 
    geom_jitter(width = 0.5, alpha = 0.3) +
    geom_smooth(method = "loess", formula = y ~ x) +
    theme_bw() +
    labs(title = "Probability of tie based on resolution",
         x = "\nResolution",
         y = "Probability\n")
}

# creates the graph: number of subjects per group vs. probability
create_subject_num_graph <- function(numTrials, numGroupsPerTrial, lowerSubNum, upperSubNum, lowerBound, upperBound, numRes, concentration, sampleType) {
  subject_num_data <- tibble(subjectNum = numeric(), prob = numeric())
  subjectNum <- c()
  prob <- c()
  for (i in lowerSubNum:upperSubNum) {
    for (x in 1:concentration) {
      subjectNum <- append(subjectNum, i)
      prob <- append(prob, calc_prob(conduct_sim(numTrials, numGroupsPerTrial, i, lowerBound, upperBound, numRes, sampleType)))
    }
  }
  subject_num_data <- subject_num_data %>% 
    add_row(subjectNum, prob) %>% 
    ggplot(mapping = aes(subjectNum, prob)) + 
    geom_jitter(width = 0.5, alpha = 0.3) + 
    geom_smooth() + 
    theme_bw() + 
    labs(title = "Probability of tie based on number of subjects",
         x = "\nNumber of Subjects in a Group",
         y = "Probability\n")
  return(subject_num_data)
}

# creates the 3D-graph: resolution vs. number of subjects per group vs. probability
create_prob_graph_3d <- function(numTrials, numGroupsPerTrial, lowerSubNum, upperSubNum, lowerNumRes, upperNumRes, lowerBound, upperBound, sampleType) {
  three_dim_data <- tibble(numRes = numeric(), subjectNum = numeric(), prob = numeric())
  numRes <- c()
  subjectNum <- c()
  prob <- c()
  for (i in lowerNumRes:upperNumRes) {
    for (j in lowerSubNum:upperSubNum) {
      numRes <- append(numRes, i)
      subjectNum <- append(subjectNum, j)
      prob <- append(prob, calc_prob(conduct_sim(numTrials, numGroupsPerTrial, j, lowerBound, upperBound, i, sampleType)))
    }
  }
  three_dim_data <- three_dim_data %>% 
    add_row(numRes, subjectNum, prob)
  
  m <- lm(prob ~ numRes + subjectNum, data = three_dim_data)
  
  numRes <- unique(three_dim_data$numRes)
  subjectNum <- unique(three_dim_data$subjectNum)
  grid <- with(three_dim_data, expand.grid(numRes, subjectNum))
  d <- setNames(data.frame(grid), c("numRes", "subjectNum"))
  vals <- predict(m, newdata = d)
  
  m <- matrix(vals, nrow = length(unique(d$numRes)), ncol = length(unique(d$subjectNum)))
  
  three_dim_data <- three_dim_data %>% 
    plot_ly(x = ~numRes, y = ~subjectNum, z = ~prob) %>% 
    add_markers() %>% 
    add_surface(x = ~numRes, y = ~subjectNum, z = ~m) %>% 
    layout(scene = list(xaxis = list(title = "Resolution"), yaxis = list(title = "Number of Subjects"), zaxis = list(title = "Probability")))
  return(three_dim_data)
}

# example call: 1000 trials, 2 groups per trial, each group has 100 subjects, uniform distribution is of range 50-250, the resolution range we are observing is between 2-100, we will be using 50 points at each resolution value in order to get a more concentrated graph
# create_prob_graph(1000, 2, 100, 50, 250, 2, 100, 50, "uniform")

```


## Coin Flip Example

y = ax^b, where y is the probability and x is the number of flips
log(y) = log(a) + b log(x)
log(y) = log(a) - 0.5 log(x)
a is function of number of faces, due to a being different
a = f(N), where N is number of faces
y = f(N) x^b
log(y) = log(f(N)) + b log(x)
log(y) = log(f(N)) - 0.5 log(x)


log(f(2)) = -0.6442
f(2) = 0.5250824          0.5?

log(f(3)) = -1.0954
f(3) = 0.3344058          0.33?

log(f(5)) = -1.6275
f(5) = 0.19642            0.2?

log(f(7)) = -1.9511
f(7) = 0.1421177          0.14?

log(f(10)) = -2.3309
f(10) = 0.09720822        0.1?

log(f(15)) = -2.7524
f(15) = 0.06377462        0.067?

f(N) = 1/N?

log(y) = log(1/N) - 0.5 log(x)
y = 1/N / x^0.5
y = 1 / (N * x^0.5)

faces,flips,probOfTie

2,1,0.499
1 / (2 * 1^0.5) = 0.5

2,22,0.12
1 / (2 * 22^0.5) = 0.107

3,10,0.1076
1 / (3 * 10^0.5) = 0.105


x = number of flips
N = number of faces
a = number of one specific (success)

x sum a = 0    x C a    (1 / N) ^ a   ((N - 1) / N) ^ (x - a)


```{r, include=TRUE, echo=FALSE, cache=TRUE}

commented_code <- function() {
  
  # coin_flip_graph <- function(numTrials, lowerFlipNum, upperFlipNum) {
#   first_matrix <- matrix(nrow = numTrials, ncol = upperFlipNum - lowerFlipNum + 1, dimnames = list(c(1:numTrials), c(lowerFlipNum:upperFlipNum)))
#   second_matrix <- matrix(nrow = numTrials, ncol = upperFlipNum - lowerFlipNum + 1, dimnames = list(c(1:numTrials), c(lowerFlipNum:upperFlipNum)))
#   for (i in lowerFlipNum:upperFlipNum) {
#     first_matrix[,i - lowerFlipNum + 1] = rbinom(numTrials, i, 0.5)
#     second_matrix[,i - lowerFlipNum + 1] = rbinom(numTrials, i, 0.5)
#   }
#   combined_matrix <- first_matrix == second_matrix
#   prob_tibble <- tibble(numFlips = numeric(), prob = numeric())
#   numFlips <- c()
#   prob <- c()
#   counts <- colCounts(combined_matrix, value = TRUE)
#   firstCount <- lowerFlipNum
#   for (count in counts) {
#     numFlips <- append(numFlips, firstCount)
#     prob <- append(prob, count / numTrials)
#     firstCount <- firstCount + 1
#   }
#   prob_graph <- prob_tibble %>% 
#     add_row(numFlips, prob) %>% 
#     ggplot(mapping = aes(numFlips, prob)) + 
#     geom_point() + 
#     theme_bw() + 
#     labs(title = "Probability of Two Groups Getting the Same Number of Heads for a Certain Number of Flips",
#          x = "\nNumber of Flips",
#          y = "Probability of Tie\n")
#   return(prob_graph)
# }
# 
# # this one has a rbinom() so it probably is incorrect for that reason
# multi_face_coin_flip_graph_vec <- function(numOutcomesVec, numTrials, flipNums) {
#   prob_tibble <- tibble(numCoinFaces = factor(), numCoinFlips = numeric(), probOfTie = numeric())
#   numCoinFaces <- c()
#   numCoinFlips <- c()
#   probOfTie <- c()
#   for (a in numOutcomesVec) {
#     first_matrix <- matrix(nrow = numTrials, ncol = length(flipNums), dimnames = list(c(1:numTrials), flipNums))
#     second_matrix <- matrix(nrow = numTrials, ncol = length(flipNums), dimnames = list(c(1:numTrials), flipNums))
#     temp <- 1
#     for (i in flipNums) {
#       first_matrix[,temp] = rbinom(numTrials, i, 0.5)
#       second_matrix[,temp] = rbinom(numTrials, i, 0.5)
#       temp <- temp + 1
#     }
#     combined_matrix <- first_matrix == second_matrix
#     counts <- colCounts(combined_matrix, value = TRUE)
#     firstCount <- 1
#     for (count in counts) {
#       numCoinFaces <- append(numCoinFaces, as.factor(a))
#       numCoinFlips <- append(numCoinFlips, flipNums[firstCount])
#       probOfTie <- append(probOfTie, count / numTrials)
#       
#       firstCount <- firstCount + 1
#     }
#   }
#   
#   prob_graph <- prob_tibble %>% 
#     add_row(numCoinFaces, numCoinFlips, probOfTie) %>% 
#     ggplot(mapping = aes(numCoinFlips, probOfTie, color = numCoinFaces)) + 
#     geom_point() + 
#     scale_x_log10() +
#     theme_bw() + 
#     labs(title = "Probability of Two Groups Getting the Same Number of Heads for a Certain Number of Flips",
#          x = "\nNumber of Flips",
#          y = "Probability of Tie\n")
#   return(prob_graph)
# }
# 
# multi_face_coin_flip_graph <- function(numOutcomesVec, numTrials, lowerFlipNum, upperFlipNum) {
#   prob_tibble <- tibble(numFlips = numeric(), prob = numeric(), numCoinFaces = factor())
#   numFlips <- c()
#   prob <- c()
#   numCoinFaces <- c()
#   for (numOutcomes in numOutcomesVec) {
#     first_matrix <- matrix(nrow = numTrials, ncol = upperFlipNum - lowerFlipNum + 1, dimnames = list(c(1:numTrials), c(lowerFlipNum:upperFlipNum)))
#     second_matrix <- matrix(nrow = numTrials, ncol = upperFlipNum - lowerFlipNum + 1, dimnames = list(c(1:numTrials), c(lowerFlipNum:upperFlipNum)))
#     for (i in lowerFlipNum:upperFlipNum) {
#       first_avg_list <- c()
#       second_avg_list <- c()
#       for (z in 1:numTrials) {
#         first_avg_list <- append(first_avg_list, signif(sum(sample(c(1:numOutcomes), i, replace=TRUE)) / i, digits=3))
#         second_avg_list <- append(second_avg_list, signif(sum(sample(c(1:numOutcomes), i, replace=TRUE)) / i, digits=3))
#       }
#       first_matrix[,i - lowerFlipNum + 1] <- first_avg_list
#       second_matrix[,i - lowerFlipNum + 1] <- second_avg_list
#     }
#     combined_matrix <- first_matrix == second_matrix
#     counts <- colCounts(combined_matrix, value = TRUE)
#     firstCount <- lowerFlipNum
#     for (count in counts) {
#       numFlips <- append(numFlips, firstCount)
#       prob <- append(prob, count / numTrials)
#       numCoinFaces <- append(numCoinFaces, as.factor(numOutcomes))
#       firstCount <- firstCount + 1
#     }
#   }
#   prob_graph <- prob_tibble %>%
#     add_row(numFlips, prob, numCoinFaces) %>%
#     ggplot(mapping = aes(numFlips, prob, color=numCoinFaces)) +
#     geom_point() +
#     # geom_smooth(method = "loess", formula = y ~ x) +
#     theme_bw() +
#     labs(title = "Probability of Two Groups Getting the Same Average Outcome for a Certain Number of Flips",
#          x = "\nNumber of Flips",
#          y = "Probability of Tie\n")
#   return(prob_graph)
# }
# 
# binom_pdf <- function(numTrials, numSuccesses, successProb) {
#   calcCombination <- factorial(numTrials) / (factorial(numSuccesses) * factorial(numTrials - numSuccesses))
#   calcProbSuccess <- successProb ^ numSuccesses
#   calcProbFailure <- (1 - successProb) ^ (numTrials - numSuccesses)
#   return(calcCombination * calcProbSuccess * calcProbFailure)
# }
# 
# binom_plot <- function(numTrials) {
#   plot <- ggplot(data.frame(x = c(1:numTrials)), aes = (x = x)) + 
#     stat_function(fun = binom_pdf(numTrials, x, 0.5))
#   return(plot)
# }
  
}

multi_face_coin_flip <- function(facesVector, numTrials, lowerFlipNum, upperFlipNum) {
  probTibble <- tibble(numCoinFaces = factor(), numCoinFlips = numeric(), probOfTie = numeric())
  numCoinFaces <- c()
  numCoinFlips <- c()
  probOfTie <- c()
  for (a in facesVector) {
    for (i in lowerFlipNum:upperFlipNum) {
      numCoinFaces <- append(numCoinFaces, as.factor(a))
      numCoinFlips <- append(numCoinFlips, i)
      probOfTie <- append(probOfTie, multi_face_prob(a, numTrials, i))
    }
  }
  graph <- probTibble %>%
    add_row(numCoinFaces, numCoinFlips, probOfTie) %>%
    ggplot(mapping = aes(numCoinFlips, probOfTie, color = numCoinFaces)) +
    geom_point() +
    theme_bw() +
    labs(title = "Probability of Two Groups Getting the Same Average Based on the Number of Outcomes",
         x = "\nNumber of Flips",
         y = "Probability of Tie\n")
  return(graph)
}

multi_face_coin_flip_vec <- function(facesVector, numTrials, flipNums) {
  probTibble <- tibble(numCoinFaces = factor(), numCoinFlips = numeric(), probOfTie = numeric())
  numCoinFaces <- c()
  numCoinFlips <- c()
  probOfTie <- c()
  for (a in facesVector) {
    for (flipNum in flipNums) {
      numCoinFaces <- append(numCoinFaces, as.factor(a))
      numCoinFlips <- append(numCoinFlips, flipNum)
      probOfTie <- append(probOfTie, multi_face_prob(a, numTrials, flipNum))
    }
  }
  graph <- probTibble %>%
    add_row(numCoinFaces, numCoinFlips, probOfTie) %>%
    ggplot(mapping = aes(numCoinFlips, probOfTie, color = numCoinFaces)) +
    geom_point() +
    theme_bw() +
    labs(title = "Probability of Two Groups Getting the Same Average Based on the Number of Outcomes",
         x = "\nNumber of Flips",
         y = "Probability of Tie\n")
  return(graph)
}

multi_face_prob <- function(numFaces, numTrials, numFlips) {
  firstGroup <- matrix(sample(1:numFaces, size = numTrials * numFlips, replace = TRUE), nrow = numTrials, ncol = numFlips)
  firstGroupMeans <- apply(firstGroup, 1, mean)
  secondGroup <- matrix(sample(1:numFaces, size = numTrials * numFlips, replace = TRUE), nrow = numTrials, ncol = numFlips)
  secondGroupMeans <- apply(secondGroup, 1, mean)
  comparisonMeans <- firstGroupMeans == secondGroupMeans
  return(mean(comparisonMeans))
}

create_coin_flip_csv <- function(numTrials, facesVector, flipsVector, csvName) {
  coin_flip_tibble <- tibble(faces = numeric(), flips = numeric(), probOfTie = numeric())
  faces <- c()
  flips <- c()
  probOfTie <- c()
  for (numFaces in facesVector) {
    for (numFlips in flipsVector) {
      flips <- append(flips, numFlips)
      probOfTie <- append(probOfTie, multi_face_prob(numFaces, numTrials, numFlips))
      faces <- append(faces, numFaces)
    }
  }
  coin_flip_tibble <- coin_flip_tibble %>% 
    add_row(faces, flips, probOfTie)
  write_csv(coin_flip_tibble, csvName)
}

analyze_trends <- function(csvName) {
  rawTibble <- read_csv(csvName)
  modifiedTibble <- rawTibble %>% 
    mutate(faces = as.factor(faces))
  modifiedTibble %>%
    ggplot(mapping = aes(log(flips), log(probOfTie), color = faces)) +
    geom_point() +
    facet_grid(~faces) + 
    theme_bw() + 
    labs(title = "Analzing Trends in Coin Flip Scenario",
         x = "\nNumber of Flips (Log)",
         y = "Probability of Tie (Log)\n")
  # lm(data = rawTibble, subset = faces == 15, log(probOfTie) ~ log(flips))
}

analyze_trends("coin_flip_data2.csv")

# plot(coin_flip_graph(80000, 2, 100))

# xfun::cache_rds({
#   
#   # write your time-consuming code in this expression
  # multi_face_coin_flip(c(2, 3, 5, 10), 10000, 1, 100)
# 
# })
# 
# xfun::cache_rds({
#   
#   # write your time-consuming code in this expression
  # multi_face_coin_flip_vec(c(2, 3, 5, 10), 10000, c(seq(100, 1000, 100)))
# 
# })

# multi_face_coin_flip_graph_vec(c(2, 3, 5, 10), 900000, c(seq(100, 1000, 100)))
# multi_face_coin_flip_graph_vec(c(2, 3, 5, 10), 900000, c(1000, 10000, 100000, 1000000))

# create_coin_flip_csv(10000, c(2, 3, 5, 10), c(1:100, 1000, 10000, 100000, 1000000), "coin_flip_data2.csv")
# create_coin_flip_csv(10000, c(2, 3, 5, 7, 10, 15), c(1:100), "coin_flip_data2.csv")
# USE THE MULTI FACE PROB FUNCTION ---- MAKES IT MUCH MUCH QUICKER ---- ADD IT TO THE MULTI FACE VEC FUNCTION
# THIS COULD BE COMPLETELY WRONG UNFORTUNATELY
# FOR NOW MAKE DO WITH LOW FLIP NUMBERS, BUT MORE VARIETY OF FACES
# USE LARGE FLIP NUMS FOR GRAPH ONLY MAYBE??

# multi_face_coin_flip_graph(c(2), 10000, 1, 100)
# coin_flip_graph(8000, 2, 100)
# coin_flip_graph2(8000, c(2:100, 1000, 10000))
```


```{r, include=TRUE, echo=FALSE, cache=TRUE}

multi_face_coin_flip(c(2, 3, 5, 10), 10000, 1, 100)

```


```{r, include=TRUE, echo=FALSE, cache=TRUE}

multi_face_coin_flip_vec(c(2, 3, 5, 10), 10000, c(seq(100, 1000, 100)))

```


```{r, include=TRUE, echo=FALSE, cache=TRUE}

# multi_face_coin_flip_vec(c(2, 3, 5, 10), 10000, c(1000, 10000, 100000, 1000000))

```


## Uniform Distribution (Resolution vs. Probability)

```{r, include=TRUE, echo=FALSE}

# plot(create_prob_graph(600, 2, 40, 50, 250, 2, 30, 1, "uniform"))

```

## Normal Distribution (Resolution vs. Probability)

```{r, include=TRUE, echo=FALSE}

# plot(create_prob_graph(600, 2, 40, 50, 250, 2, 30, 1, "normal"))

```

## Geometric Distribution (Resolution vs. Probability)

```{r, include=TRUE, echo=FALSE}

# plot(create_prob_graph(600, 2, 40, 50, 250, 2, 30, 1, "geometric"))

```

## Number of Subjects in a Group (Number of Subjects Per Group vs. Probability)

```{r, include=TRUE, echo=FALSE}

# plot(create_subject_num_graph(600, 2, 20, 60, 50, 250, 5, 1, "uniform"))

```

## Resolution vs. Number of Subjects Per Group vs. Probability

```{r, include=TRUE, echo=FALSE}

# create_prob_graph_3d(100, 2, 20, 50, 5, 20, 50, 250, "uniform")

```





